---
title: "Using a manually downloaded CIMIS CSV with TrackTrap"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using a manually downloaded CIMIS CSV with TrackTrap}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

## When to use this

Use `weather_source = "cimis_csv"` when you intend to use station data downloaded manually instead
of a gridded product (Daymet/Open-Meteo), or when you are working offline and have the pre-downloaded data.

## 1. Download the report

At the [CIMIS website](https://cimis.water.ca.gov): **Reports** > select
your station > report type **Daily** > units **English (°F)** > export as
CSV. The file will include `Date`, `Max Air Temp (F)`, and `Min Air Temp (F)`,
along with other columns that `calc_pest_phenology()` ignores.

## 2. Know how `read.csv()` renames columns

`read.csv()` processes headers that contain spaces or parentheses:
`Max Air Temp (F)` becomes `Max.Air.Temp..F.`, and `Min Air Temp (F)` becomes
`Min.Air.Temp..F.` — exactly the names `calc_pest_phenology()` expects.

```{r, eval = TRUE}
example_path <- system.file("/extdata", "example_cimis_daily.csv", package = "TrackTrap")
names(read.csv(example_path))
```

## 3. Confirm the date format

CIMIS exports dates as `MM/DD/YYYY`, which `calc_pest_phenology()` accepts
directly — no manual conversion needed. Date should be
input as `M/D/YYYY` or `MM/DD/YYYY` character strings (e.g. `"4/1/2024"`),
not as pre-converted `Date` objects. Any date that doesn't match this format
will cause the function to stop and name the offending value.

## 4. Run `calc_pest_phenology()`

Set `weather_source = "cimis_csv"` exactly (not `"cimis"`).

`calc_pest_phenology()` does **not** read station location from the CSV — it
only extracts `Date`, `Max.Air.Temp..F.`, and `Min.Air.Temp..F.` from the file.
`lat`/`lon` arguments passed to `calc_pest_phenology()` are ignored in this
mode (a message is printed to confirm this). If `lat`/`lon` are omitted, the
result's `lat`/`lon` attributes remain `NULL`, and `plot_trap_phenology()`
will omit the location title unless coordinates are provided directly to
that plotting call instead (see Step 5).

```{r, eval = TRUE}
library(TrackTrap)

trap_df <- data.frame(
  date = c("4/1/2024", "4/8/2024", "4/15/2024", "4/22/2024"),
  trap_counts = c(0, 0, 3, 9)
)

result <- calc_pest_phenology(
  trap_df, pest = "OLFF",
  weather_source = "cimis_csv",
  cimis_csv_path = example_path
)
```

## 5. Plot it

Because location is not carried through from the CIMIS CSV, provide `lat` and
`lon` explicitly to `plot_trap_phenology()` if you want the station location
shown in the plot title.

```{r, eval = TRUE, fig.width = 7, fig.height = 5}
plot_trap_phenology(result, lat = 36.3302, lon = -119.2921)
```